### 2025-03-07 paper/markdown/rebuttal_letter.Rmd
DLCO <80% predictions and DLCO values
Analysis of predictions of DLCO < 80% in observations with increasing DLCO values expressed as percentages of the patient's reference. Drop-out rates and frequency of LFT and CT abnormalities included in the manuscript text.

---
title: "Artificial intelligence-assisted analysis of CT abnormalities during COVID-19 recovery"
subtitle: "Rebuttal Letter"
author: "Department of Radiology, Medical University of Innsbruck"
date: "`r format(Sys.time(), '%Y-%m-%d')`"

output: 
  bookdown::html_document2:
    css: "style.css"

bibliography: ct.bib

csl: frontiers_medical.csl

header-includes:
  \usepackage{longtable}
  \usepackage{tabu}
  \usepackage{caption}
  \usepackage{makecell}
  \usepackage{pdflscape}
  \usepackage{array}
  \usepackage{booktabs}
  \usepackage{threeparttable}
  \usepackage{threeparttablex}
  \usepackage{wrapfig}
  \usepackage{multirow}
  \usepackage[normalem]{ulem}
  \usepackage{colortbl}
  \usepackage{xcolor}
  \usepackage{float} \floatplacement{figure}{H} \floatplacement{table}{H}

---

```{r, setup, include = FALSE}

library(Cairo)

opts_chunk$set(echo = FALSE, 
                      warning = FALSE, 
                      message = FALSE, 
                      dev = "CairoPNG", 
                      dpi = 600)

set_flextable_defaults(font.family = "Cambria", 
                       font.size = 10)

```

\newpage

# Reviewer 1

## Point 1

### Issue

Add results in the abstract as well.

### Response

## Point 2

### Issue

The introduction is not sufficient.

### Response

## Point 3

### Issue

The literature review part needs to be added.

### Response

## Point 4

### Issue

After adding the literature review, add a summary table and clearly mention this study's research gaps and contributions.

### Response

## Point 5

### Issue

More details, such as features, are required in the study data.

### Response

## Point 6

### Issue

The paper lacks details about hyper-parameters for machine learning models. 
For example, what was the C and class weight value for the SVM model?

### Response

## Point 7

### Issue

Have you considered the GridSearchCV for model hyper-parameters or set it manually?

### Response

## Point 8

### Issue

Comparison with state-of-the-art studies is missing.

### Response

## Point 9

### Issue

The Practical Implication of the study is missing.

### Response

## Point 10

### Issue

Conclusion also not well presented.

### Response

# Reviewer 2

## Point 11

### Issue

The introduction does not sufficiently contextualize how other published multi-parameter or machine learning approaches have succeeded or failed. What specific gaps remain? A more detailed table is needed to perform the literature review.

### Response

## Point 12

### Issue

What is new or improved about your multi-parameter modeling approach compared to existing frameworks?

### Response

## Point 13

### Issue

More details about selection and potential biases (e.g., dropouts, missing visits) are needed for the data set.

### Response

## Point 14

### Issue

It remains unclear how each scoring system aligns, differs, or complements the other.

### Response

Overemphasis on “satisfactory accuracy” without deeper clinical context.

## Point 15

### Issue

The neural network model performed poorly for DLCO.

### Response

## Point 16

### Issue

The SHAP analysis identifies “top 15” predictors. However, the manuscript does not address correlations among those predictors or discuss whether certain features might be redundant.

### Response

## Point 17

### Issue

With “opacity cutoff: 0.12%” or “high opacity cutoff: 0.002%” appear extremely small. The clinical significance of such minimal changes in lung volume being “abnormal” is unclear—are these thresholds truly meaningful in practice?

### Response

## Point 18

### Issue

The paper concedes “model overfitting” could be an issue. However, it lacks any systematic attempt (e.g., a learning-curve analysis or thorough regularization strategy) to demonstrate that overfitting is minimal or under control.

### Response

# References